Purpose

This script automates the process of loading, processing, and merging CSV data files across multiple years. It is designed to standardize and filter data based on predefined variables for further analysis.

Quick Start Guide

1. Data Preparation
	•	Place CSV files in the folder: data/strukturerhebung/.
	•	Ensure file names follow the pattern: {year}_SE_RS_CSV/{file_type}{year}.csv.
	•	Example for household data: 2011_SE_RS_CSV/hhm_2011.csv.

2. Run the Script

Step 1: Load Variables
	1.	Predefined Variables:
	•	variables_zp.
	•	variables_hh.
	•	variables_hhm.
	2.	Load these variables into the script from previously saved .rds files:
		path_save <- "questions_answers/home_owner_ship/"
		variables_zp <- readRDS(paste0(path_save, "variables_zp.rds"))
		variables_hh <- readRDS(paste0(path_save, "variables_hh.rds"))
		variables_hhm <- readRDS(paste0(path_save,"variables_hhm.rds"))
Step 2: Merge Data Across Years
	•	Use the merging() function to combine data across multiple years:
		merged_data <- merging(2010, 2021, "zp_", variables_zp)
		  *Change parameters to csv file that you need
	•	The merged dataset will be saved as a .csv file in the **path_save** directory.

3. Output Files
	•	Filtered Data:
	•	Individual-year data are saved temporarily in R objects for inspection.
	•	Merged Data:
	•	Combined data across years is saved as zp_merged_data.csv, hh_merged_data.csv or hhm_merged_data.csv in the output directory.

```{r}
library(dplyr)
source("Renaming_Columns.R")
```

```{r}

path_save = "Report/data/"
variables_zp <- readRDS(paste0(path_save,"variables_zp.rds"))
variables_hh <- readRDS(paste0(path_save,"variables_hh.rds"))
# variables_hhm <- readRDS(paste0(path_save,"variables_hhm.rds"))
```

```{r}
# merged_variables <- c(vatriables_zp,variables_hh)
```

```{r}

load_data <- function(year, file_type, variables) {
  file_path <- paste0("data/strukturerhebung/", year, "_SE_RS_CSV/", file_type, year, ".csv")
  
  if (file.exists(file_path)) {
    if (year == 2011 & file_type == "hhm_"){
      data <- read.csv(file_path, sep = ",")
    }else{
    data <- read.csv(file_path, sep = ";")
    }
    colnames(data) <- toupper(colnames(data))
    data$YEAR_QUIZ <- year
    data <- data %>% select(any_of(c(variables, "YEAR_QUIZ")))
    return(data)
  } else {
    message(paste("File does not exist:", file_path))
    return(NULL)
  }
}
```

```{r}

load_all_data <- function(start_year, end_year, file_type, variables) {
  data_list <- list()
  
  for (year in start_year:end_year) {
    data <- load_data(year, file_type, variables)
    if (!is.null(data)) {
      data_list[[as.character(year)]] <- data
    }
  }
  
  return(data_list)
}
```

```{r}

unify_column_types <- function(data_list) {
  all_columns <- unique(unlist(lapply(data_list, names)))
  
  for (col in all_columns) {
    column_types <- sapply(data_list, function(df) {
      if (col %in% names(df)) class(df[[col]]) else NA
    })
    
    if (length(unique(na.omit(column_types))) > 1) {
      data_list <- lapply(data_list, function(df) {
        if (col %in% names(df)) df[[col]] <- as.character(df[[col]])
        return(df)
      })
    }
  }
  
  return(data_list)
}
```

```{r}
merging <- function(start_year, end_year, file_type, variables, path_save){
  data_list <- load_all_data(start_year, end_year, 
                             file_type, variables)
  
  rename_mappings <- list(
    "SECCITIZ" = "SECCITIZENSHIP",
    "WORKSTATUSDETAIL" = "WORKSTATUS_DETAIL"
  )
  
  data_list <- rename_columns(data_list, rename_mappings)
  data_list <- unify_column_types(data_list)
  
  
  merged_data <- data_list[[as.character(start_year)]]
  for (year in (start_year + 1):end_year) {
    if (!is.null(data_list[[as.character(year)]])) {
      merged_data <- bind_rows(merged_data, data_list[[as.character(year)]])
    }
  }
  
  write.csv(merged_data, paste0(path_save, file_type, "merged_data",".csv"), row.names = FALSE)
  
  return(merged_data)
}
```

```{r}
zp_data_list <- merging(2010, 2021, "zp_", variables_zp, path_save)
```

```{r}
hh_data_list <- merging(2010, 2021, "hh_", variables_hh, path_save)
```


```{r}
# hhm_data_list <- merging(2010, 2021, "hhm_", c("HOUSEHOLDYEARLYID",variables_hhm))
```
